Donor Preferences in Canadian Political Financing

### 2024-12-14 paper/paper.qmd
finalized idealized survey

---
title: "Donor Preferences in Canadian Political Financing"
subtitle: "Understanding Donor Support for Ruling vs. Opposition Parties Between 2013 - 2024"
author: 
  - Maria Mangru
thanks: "Code and data are available at: [https://github.com/MariaMangru/Ontario_Political_Donors](https://github.com/MariaMangru/Ontario_Political_Donors)."
date: today
date-format: long
abstract: "This study investigates how the ruling political parties in Canada's federal and Ontario provincial governments affect the donation behaviors of individual contributors. By analyzing donation records from 2013 to 2024, the research explores whether a party being in power influences the total amount of financial support it receives. The findings reveal that opposition parties often receive more donations than those in power, highlighting donor preferences to support challengers. This insight enhances our understanding of political financing and can help parties and policymakers develop more effective fundraising strategies."
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

@@

## Financial influence shapes political landscapes

@@

## International Comparisons and the Importance of Regulation
These trends are not confined to Canada. In the United Kingdom, for example, large private donations, often from very wealthy “superdonors”, widen the gap between major parties and smaller competitors [@draca2022financing]. Such patterns highlight a broader democratic challenge, as money can concentrate power and influence political direction in ways that may not always reflect broad public preferences. Rather than serving purely as a resource that encourages healthy competition, political donations can also become a tool that benefits certain parties and voices more than others.
Regulation and transparency are essential. Clearer rules and better public reporting help citizens understand who funds whom and why. Drawing lessons from other countries may guide reforms aimed at limiting undue influence. Stronger oversight, spending caps, and easy-to-access donor databases can empower voters to make informed choices. Without these safeguards, the already powerful, whether individuals, industries, or interest groups, can gain even more sway, risking an erosion of trust in democratic institutions.

@@

##Idealized Methodology

@@

-   Details: Expenses related to accessing online panel respondents,
    purchasing targeted advertising on social media, and collaborating
    with panel providers. Funds also cover the services of statisticians
    to ensure the sampling aligns with the population distribution.

@@

-   Details: Allocated for compensating participants with \$10
    electronic gift cards or funding the national sweepstakes prizes.
    Incentives are crucial for improving response rates and attracting a
    diverse respondent pool.

@@

-   Details: Operational costs for administering the survey, including
    online platform fees, telephone survey expenses, interviewer
    training, and call center operations. This ensures data collection
    is efficient and reaches a broad audience.

@@

-   Details: Funds dedicated to analyzing the collected data, including
    hiring data analysts proficient in statistical software, conducting
    comprehensive analyses, and preparing detailed reports and
    visualizations to communicate findings effectively.

@@

-   Details: Covers the costs of obtaining ethical approvals, ensuring
    compliance with privacy legislation (e.g., Personal Information
    Protection and Electronic Documents Act), and implementing robust
    data security measures to protect respondent information.

@@

This survey helps to gather data about those who donate to political
parties and acquire more qualitative details about a sample of the
population.

The survey can be accessed using [this
link](https://docs.google.com/forms/d/e/1FAIpQLSdj2RxS9fjZ_n6Q54zjSyoJ7tAJT7ssuTF3V1tVk6a_Rhyp4w/viewform).

@@

@@

Age: Multiple-choice options

@@

@@

Income Level:

@@

Education Level: Highest degree obtained

@@

@@

@@

Are you currently registered to donate to political parties?

@@

Average Donation Amount:

@@

Preferred Donation Channels:

-   Online platforms

@@

-   In-person events

@@

@@

-   Candidate's leadership qualities

-   Party's stance on key issues

-   Party's past performance

-   Recommendations from peers or influencers

-   Media coverage

@@

Have recent political events influenced your donation behavior?

@@

If yes, please specify the events and how they influenced your
donations. (Open-ended)

How do you perceive the effectiveness of political campaigns in
encouraging donations?

@@

@@

What could political parties do to encourage more donations from
supporters? (Open-ended) -

Any additional comments or suggestions regarding political donations?
(Open-ended)

>>>>

---
title: "Donor Preferences in Canadian Political Financing"
subtitle: "Understanding Donor Support for Ruling vs. Opposition Parties Between 2013 - 2024"
author: 
  - Maria Mangru
thanks: "Code and data are available at: [https://github.com/MariaMangru/Ontario_Political_Donors](https://github.com/MariaMangru/Ontario_Political_Donors)."
date: today
date-format: long
abstract: "This study investigates how the ruling political parties in Canada's federal and Ontario provincial governments affect the donation behaviors of individual contributors. By analyzing donation records from 2013 to 2024, the research explores whether a party being in power influences the total amount of financial support it receives. The findings reveal that opposition parties often receive more donations than those in power, highlighting donor preferences to support challengers. This insight enhances our understanding of political financing and can help parties and policymakers develop more effective fundraising strategies."
format: 
  pdf:
    toc: true
    toc-depth: 3
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

@@

## Financial Influence Shapes Political Landscapes

@@

## International Comparisons and the Importance of Regulation

These trends are not confined to Canada. In the United Kingdom, for
example, large private donations, often from very wealthy “superdonors”,
widen the gap between major parties and smaller competitors
[@draca2022financing]. Such patterns highlight a broader democratic
challenge, as money can concentrate power and influence political
direction in ways that may not always reflect broad public preferences.
Rather than serving purely as a resource that encourages healthy
competition, political donations can also become a tool that benefits
certain parties and voices more than others. Regulation and transparency
are essential. Clearer rules and better public reporting help citizens
understand who funds whom and why. Drawing lessons from other countries
may guide reforms aimed at limiting undue influence. Stronger oversight,
spending caps, and easy-to-access donor databases can empower voters to
make informed choices. Without these safeguards, the already powerful,
whether individuals, industries, or interest groups, can gain even more
sway, risking an erosion of trust in democratic institutions.

@@

## Appendix C: Idealized Methodology

@@

    -   Expenses related to accessing online panel respondents,
        purchasing targeted advertising on social media, and
        collaborating with panel providers. Funds also cover the
        services of statisticians to ensure the sampling aligns with the
        population distribution.

@@

    -   Allocated for compensating participants with \$10 electronic
        gift cards or funding the national sweepstakes prizes.
        Incentives are crucial for improving response rates and
        attracting a diverse respondent pool.

@@

    -   Operational costs for administering the survey, including online
        platform fees, telephone survey expenses, interviewer training,
        and call center operations. This ensures data collection is
        efficient and reaches a broad audience.

@@

    -   Funds dedicated to analyzing the collected data, including
        hiring data analysts proficient in statistical software,
        conducting comprehensive analyses, and preparing detailed
        reports and visualizations to communicate findings effectively.

@@

    -   Covers the costs of obtaining ethical approvals, ensuring
        compliance with privacy legislation (e.g., Personal Information
        Protection and Electronic Documents Act), and implementing
        robust data security measures to protect respondent information.

@@

This survey helps to gather data about those who donate to political
parties and acquire more qualitative details about a sample of the
population. The survey can be accessed using [this
link](https://docs.google.com/forms/d/e/1FAIpQLSdj2RxS9fjZ_n6Q54zjSyoJ7tAJT7ssuTF3V1tVk6a_Rhyp4w/viewform).

@@

**Introduction**

Thank you for participating in this important survey. Your responses
will help us understand how Canadians make decisions about donating to
political parties at both the federal and Ontario provincial levels. By
sharing your perspectives, you will contribute to a deeper understanding
of what motivates political donations, what factors shape donor
decisions, and how parties can improve their engagement with supporters.
Your answers are completely confidential and will be used solely for
research purposes.

@@

Age:

@@

-   65+

@@

Annual Household Income Level:

@@

-   Prefer not to say

Highest Level of Education Completed:

@@

-   Green

@@

-   No specific affiliation

@@

Have you ever donated to a political party?

-   Yes, currently donate

-   Yes, but not currently

-   No, never donated

@@

Average Annual Donation Amount (to all political parties combined):

@@

-   Prefer not to say

Preferred Donation Channels: (Select all that apply)

-   Online platforms (e.g., party websites, crowdfunding sites)

@@

-   In-person events (e.g., fundraisers, rallies)

-   Phone solicitations

@@

-   Personal connection to a candidate

@@

-   Candidate’s leadership qualities

-   Party’s stance on key issues

-   Party’s track record and past performance

-   Recommendations from peers, family, or community leaders

-   Media coverage and public opinion polls

-   Urgency or importance of the upcoming election

@@

-   Have recent political events (e.g., elections, leadership changes,
    policy announcements) influenced your donation behavior?

@@

    If yes, please specify the event(s) and how they influenced your
    donations:\
    *(Open-ended)*

-   How do you perceive the effectiveness of political campaigns in
    encouraging donations?

@@

-   Have you ever stopped donating or reduced donations because of a
    party’s perceived performance or power status (e.g., losing seats,
    scandals)?

    -   Yes (please explain why) *(Open-ended)*

@@

-   In your opinion, what could political parties do to encourage more
    donations from supporters?\
    *(Open-ended)*

-   Any additional comments or suggestions regarding political
    donations?\
    *(Open-ended)*

**Thank you for participating!** Your feedback is valuable and will help
researchers and political parties understand the dynamics of political
giving in Canada. Your insights contribute to more informed policies and
practices, ultimately shaping a more responsive and transparent
political landscape.
